---
title: "Willamette River Salmon"
author: "Allison Hacker, Sydney Rilum, Elmera Azadpour"
date: "2/2/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(janitor)
library(here)
library(tsibble)
library(lubridate)
library(slider)
library(feasts)
library(patchwork)
```

## Overview {.tabset}

Image w/ caption

Summary: (3-4 sent)

Map of fish ladder location

Data Citation:


### Daily Time Series Data

```{r}
# Read in the data and clean up column names
salmon <- read_csv(here("data", "willamette_fish_passage.csv")) %>% 
  clean_names()
```




### Seasonplots

```{r}
# Convert salmon data set into a time series data frame (tsibble)
salmon_ts <- salmon %>% 
  mutate(date = lubridate::mdy(date)) %>%  #converts date to readable R form
  as_tsibble(key = NULL, index = date)

# Create seasonplots
coho_season <- salmon_ts %>% 
  gg_season(y = coho) +
  labs(y = "Coho",
       x = "",
       title = "Salmon Seasonplots")

jack_coho_season <- salmon_ts %>% 
  gg_season(y = jack_coho) +
  labs(y = "Jack Coho",
       x = "")

steelhead_season <- salmon_ts %>% 
  gg_season(y = steelhead) +
  labs(y = "Steelhead",
       x = "Month")

# Combine seasonplots into a single compound figure using the `patchwork` package
coho_season / jack_coho_season / steelhead_season &
  theme_light() +
  theme(legend.position = "none") 
```

**Figure 2.** Seasonplots for coho, jack coho, and steelhead salmon species counts at the Willamette Falls fish ladder on the Willamette River in Oregon. Each colored line represents count data for a single year, including the years 2001 to 2010. Data: ???

Major Trends:

- Coho and jack coho salmon are only present from late August to early November, with counts peaking in the month of October.

- Steelhead salmon are present year-round, with counts highest in the spring (from March to June) and lowest in the fall.



### Annual Counts




